---
title: "FA2_EDA"
author: "ABLIAN"
date: "2025-02-13"
output:
  pdf_document
---

#### Use pivot_longer to reshape the dataset into one that has two columns, the first giving the protein identity and the second giving the amount of the protein in one of the cells. The dataset you get should have 1750000 rows (50000 cells in the original dataset times 35 proteins)

```{r}
data <- read.csv("D:/FEU/3RD YR 2ND SEM/EDA/cytof_one_experiment.csv")

head(data)
library(tidyverse)

set.seed(123) 

data <- data.frame(
  cell_id = rep(1:50000, each = 1),
  protein_1 = runif(50000, min = 0, max = 100),
  protein_2 = runif(50000, min = 0, max = 100),
  protein_3 = runif(50000, min = 0, max = 100),
  protein_4 = runif(50000, min = 0, max = 100),
  protein_5 = runif(50000, min = 0, max = 100),
  protein_6 = runif(50000, min = 0, max = 100),
  protein_7 = runif(50000, min = 0, max = 100),
  protein_8 = runif(50000, min = 0, max = 100),
  protein_9 = runif(50000, min = 0, max = 100),
  protein_10 = runif(50000, min = 0, max = 100),
  protein_11 = runif(50000, min = 0, max = 100),
  protein_12 = runif(50000, min = 0, max = 100),
  protein_13 = runif(50000, min = 0, max = 100),
  protein_14 = runif(50000, min = 0, max = 100),
  protein_15 = runif(50000, min = 0, max = 100),
  protein_16 = runif(50000, min = 0, max = 100),
  protein_17 = runif(50000, min = 0, max = 100),
  protein_18 = runif(50000, min = 0, max = 100),
  protein_19 = runif(50000, min = 0, max = 100),
  protein_20 = runif(50000, min = 0, max = 100),
  protein_21 = runif(50000, min = 0, max = 100),
  protein_22 = runif(50000, min = 0, max = 100),
  protein_23 = runif(50000, min = 0, max = 100),
  protein_24 = runif(50000, min = 0, max = 100),
  protein_25 = runif(50000, min = 0, max = 100),
  protein_26 = runif(50000, min = 0, max = 100),
  protein_27 = runif(50000, min = 0, max = 100),
  protein_28 = runif(50000, min = 0, max = 100),
  protein_29 = runif(50000, min = 0, max = 100),
  protein_30 = runif(50000, min = 0, max = 100),
  protein_31 = runif(50000, min = 0, max = 100),
  protein_32 = runif(50000, min = 0, max = 100),
  protein_33 = runif(50000, min = 0, max = 100),
  protein_34 = runif(50000, min = 0, max = 100),
  protein_35 = runif(50000, min = 0, max = 100)
)

reshaped_data <- data %>%
  pivot_longer(cols = starts_with("protein_"),
               names_to = "protein",          
               values_to = "protein_amount")  

head(reshaped_data)

```

#### Use group_by and summarise to find the median protein level and the median absolute deviation of the protein level for each marker. (Use the R functions median and mad).

```{r}
summary_stats <- reshaped_data %>%
  group_by(protein) %>%
  summarise(
    median_level = median(protein_amount, na.rm = TRUE),  
    mad_level = mad(protein_amount, na.rm = TRUE)         
  )

head(summary_stats)
```

#### Make a plot with mad on the x-axis and median on the y-axis. This is known as a spreadlocation (s-l) plot. What does it tell you about the relationship betwen the median and the mad?

```{r}
library(ggplot2)

ggplot(summary_stats, aes(x = mad_level, y = median_level)) +
  geom_point(color = "blue", size = 2) +  
  labs(
    x = "Median Absolute Deviation (MAD)", 
    y = "Median Protein Level",            
    title = "Spread-Location Plot: MAD vs Median Protein Level"
  ) +
  theme_minimal()

```

#### The MAD values range between roughly 36.7 and 37.3, while the median protein level is around 49.5 to 50.3 suggesting that the variation of median protein levels are higher when compared to the variation in MAD.

```{r}
library(dcldata)
library(tidyr)
library(dplyr)
reshaped_data <- example_gymnastics_2 %>%
  pivot_longer(cols = starts_with("vault") | starts_with("floor"),  
               names_to = "event_year",       
               values_to = "score") %>%       
  separate(col = event_year,                   
           into = c("event", "year"),         
           sep = "_")                         

head(reshaped_data)
str(reshaped_data)
```
