---
title: "SA1"
author: "ABLIAN"
date: "2025-03-14"
output: pdf_document
---

####Univariate Data Analysis
Load the dataset and summarize its structure.

```{r}
library(ggplot2)
library(dplyr)
data <- read.csv("D:/FEU/3RD YR 2ND SEM/EDA/EDA_Ecommerce_Assessment.csv")

head(data)
```
Create histograms and boxplots to visualize the distribution of Purchase_Amount, Number_of_Items, and Satisfaction_Score.
```{r}
plot1<-ggplot()+
      geom_histogram(aes(x=Purchase_Amount), data=data, color = "black")

plot2<-ggplot()+
      geom_histogram(aes(x=Number_of_Items), data=data,bins= 9, color="black", fill="lightblue")

plot3<-ggplot()+
      geom_histogram(aes(x=Satisfaction_Score), data=data, bins=5, color="black", fill="lightpink")

print(plot1)
print(plot2)
print(plot3)
```
Compute measures of central tendency (mean, median, mode) and spread (variance, standard deviation, IQR) for Purchase_Amount.
```{r}
plot1 <- ggplot() +
  geom_boxplot(aes(y = Purchase_Amount), data = data, color = "black")

plot2 <- ggplot() +
  geom_boxplot(aes(y = Number_of_Items), data = data, color = "black", fill = "lightblue")

plot3 <- ggplot() +
  geom_boxplot(aes(y = Satisfaction_Score), data = data, color = "black", fill = "lightpink")

print(plot1)
print(plot2)
print(plot3)
```

```{r}
mean_purchase <- mean(data$Purchase_Amount)
median_purchase <- median(data$Purchase_Amount)
mode_purchase <- names(sort(-table(data$Purchase_Amount)))[1]

var_purchase <- var(data$Purchase_Amount)
sd_purchase <- sd(data$Purchase_Amount)
IQR_purchase <- IQR(data$Purchase_Amount)

summary_df <- data.frame(
  Statistic = c("Mean", "Median", "Mode", "Variance", "Standard Deviation", "Interquartile Range"),
  Value = c(mean_purchase, median_purchase, mode_purchase, var_purchase, sd_purchase, IQR_purchase)
)

print(summary_df)
```

Compare the distribution of Browsing_Time and Purchase_Amount across different Gender groups using density plots.
```{r}
ggplot(data=data, aes(x=Purchase_Amount, group=Gender, fill=Gender, alpha=0.5)) +
  geom_density(adjust=1.5) 
ggplot(data=data, aes(x=Browsing_Time, group=Gender, fill=Gender, alpha=0.5)) +
  geom_density(adjust=1.5) 
```

Apply a logarithmic or square root transformation on Browsing_Time and evaluate changes in skewness.
```{r}
library(moments)

data <- data%>%
  mutate(Log_Browsing_Time = log(Browsing_Time + 1))  # +1 to avoid log(0)

skewness_value <- skewness(data$Log_Browsing_Time, na.rm = TRUE)
print(paste("Skewness of Log_Browsing_Time:", skewness_value))
```
This is skewed to the left.


Fit a simple linear regression model predicting Purchase_Amount based on Browsing_Time. Interpret the results.
```{r}
model<-lm(Purchase_Amount~Browsing_Time, data=data)
summary(model)

```

There is no strong evidence that browsing time significantly affects purchase amounts (p > 0.05). The model suggests that when browsing time is zero, customers spend about ₱253 on average. While browsing time shows some relationship with spending, the connection isn't statistically strong enough in this data to be certain it's not just random variation.

Use ggplot2 (or equivalent) to create scatter plots and regression lines.
```{r}
res_purchase <- resid(model)
data$res_purch <- res_purchase

ggplot(data=data, aes(x=Browsing_Time, y=res_purchase)) +
  geom_point()+
  geom_smooth()

```
#### Bivariate Data Analysis
Create scatter plots to explore the relationship between Purchase_Amount and Number_of_Items.
```{r}
ggplot(data=data, aes(x=Number_of_Items, y=Purchase_Amount)) +
  geom_point()
```
Purchase_Amount is varying from 0-500. This indicates that there are products that have higher value than others and since Purchase_Amount vary across all points, there are no clear relationship between Number_of_Items and Purchase_Amount.

Fit a polynomial regression model for Purchase_Amount and Browsing_Time and compare it with a simple linear model.
```{r}
model<-lm(Purchase_Amount~poly(Browsing_Time, 2), data=data)

res_purch_poly <- resid(model)
data$res_purch_poly <- res_purch_poly

p1<-ggplot(data=data, aes(x=Browsing_Time, y=res_purch_poly)) +
      geom_point()+
      geom_smooth()

p2<-ggplot(data=data, aes(x=Browsing_Time, y=res_purch_poly)) +
      geom_point()+
      geom_smooth()

p1
p2

```
There are no difference between the two even if you look very closely.

Apply LOESS (Locally Estimated Scatterplot Smoothing) to Purchase_Amount vs. Browsing_Time and visualize the results.
```{r}
ggplot(data=data, aes(x=Browsing_Time, y=res_purch_poly)) +
  geom_point()+
  geom_smooth(method="loess", span=0.4, se=TRUE)

```
LOESS (blue squiggly line) has a few fluctuations throughout the map, but there is still no strong relationship between the two.

Compare robust regression methods (Huber or Tukey regression) with ordinary least squares (OLS).
```{r}
library(MASS)
huber_model <- rlm(Purchase_Amount ~ Browsing_Time, data = data, psi = psi.huber)
summary(huber_model)

```

```{r}
ols_model <- lm(Purchase_Amount ~ Browsing_Time, data =data)
summary(ols_model)
```
```{r}
ggplot(data, aes(x = Browsing_Time, y = Purchase_Amount)) +
  geom_point(alpha = 0.5, color = "black") +
  geom_smooth(method = "lm", color = "red", se = FALSE, linetype = "dashed") +  
  geom_abline(slope = coef(huber_model)[2], intercept = coef(huber_model)[1], color = "lightblue", size = 1.2, alpha = 0.4)  
```
Looking at the summary they have minor differences to each other but in the graph they traverse the same line and are almost the same, with minimal differences.

#### Trivariate/Hypervariate Data Analysis
Explore interaction effects between Browsing_Time and Category on Purchase_Amount using interaction plots.
```{r}
ggplot(data, aes(x = Browsing_Time, y = Purchase_Amount, color = Category)) +
  geom_smooth(method = "lm", se = FALSE) 

```
We can see that there are different interactions between purchase amount and browsing time and that there are categories like books and home & kitchen that have a negative relationship (intersecting point below) and positive relationships with clothing and beauty (intersecting point above).


Create coplots of Purchase_Amount against Browsing_Time for different levels of Category.

```{r}
ggplot(data, aes(x = Browsing_Time, y = res_purchase)) +
  geom_point(alpha = 0.4, color = "red") +
  geom_smooth(method = "lm", se = TRUE, color = "black") +
  facet_wrap(~ Category)


```


```{r}
ggplot(data, aes(x = Browsing_Time, y = Number_of_Items, z = Purchase_Amount)) +
  geom_density_2d_filled()


```
There are higher densities on the yellow region as seen in the legend to the right and most common behaviours in terms of browsing time and number of items.

Perform multiple regression with Purchase_Amount as the dependent variable and Browsing_Time, Number_of_Items, and Satisfaction_Score as predictors. Perform model selection and assess variable importance.

```{r}
model_multi<-lm(Purchase_Amount~Browsing_Time+Number_of_Items+Satisfaction_Score ,data=data)
summary(model_multi)

```


```{r}
stepwise<-step(model_multi, direction="both")
summary(stepwise)
```
We can say that it removed all the predictors leaving behind Purchase_Amount alone using the stepwise model,indicating that the 3 predictors did not have significance value in explaining Purchase_Amount.

